---
title: "Lloyd 2016 DR Birds: Add zeros when species not captured"
author: "brouwern@gmail.com"
date: "February 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Add zeros

Need to add zeros for year-Location combinations when a species was not captured BUT banding DID occur

* initial file: Lloyd_PeerJ2916_cleaned_data_2_7_2017.csv
* final file:   Lloyd_PeerJ2916_cleaned_with_0s_2_7_2017.csv

## Load Lloyd data

```{r}
setwd("C:/Users/lisanjie2/Dropbox/Scott_E_Aviary_Projects")

spp.ann.tot <- read.csv(file = "Lloyd_PeerJ2916_cleaned_data_2_7_2017.csv")
```



## Examine data

Most species were captured very rarely in any given year
```{r}
hist(spp.ann.tot$spp.ann.tot)
```






There now should be NAs inserted into the abundance column in years when a species wasn't observed at a particular site
```{r}
summary(spp.ann.tot.b$spp.ann.tot)
```


Convert those NAs to zeros
```{r}
i.NA <- which(is.na(spp.ann.tot.b$spp.ann.tot) == TRUE)
spp.ann.tot.b[i.NA,"spp.ann.tot"] <- 0
```


Check that zeros have been added
```{r}
summary(spp.ann.tot.b$spp.ann.tot)
```



# Save data

```{r}
write.csv(spp.ann.tot.b, 
          file = "Lloyd_PeerJ2916_cleaned_with_0s_2_7_2017.csv")
```

