---
title: "Lloyd 2016 DR Birds: PALO GLMM"
author: "brouwern@gmail.com"
date: "February 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# ISSUES / EDITS

* NOTE: the file that gets read it already has spp obs in < 3 years already removed
* Still seems to be some issues w/ data
* extracting predictions when...
    + offest
    + nesting random effects
    + (what did I do w/Costa Rica?)


# Introduction

This script splits off the PALO data from the full Lloyd et al 2016 dataset uses it to demonstrate the mixed modeling approach

#Data setup

## Load Lloyd data

```{r}
setwd("C:/Users/lisanjie2/Dropbox/Scott_E_Aviary_Projects")

dat <- read.csv(file = "Lloyd_PeerJ2916_for_regression_2_16_2017.csv")
```



## Load spp info

* skip = 7 skipst he 7 rows of header info in this fiel
```{r}
spp.info <- read.csv(file = "Latta_Lloyd_species_list_SLvs1.csv",
                     skip = 7)

head(spp.info)

```



## Merge

```{r}
names(dat)[3] <-"spp.code"

dat2 <- merge(dat, spp.info)

dim(dat2)
dim(dat)

dat <- dat2


#remove some extra rows
dat <- dat[,-c(2,4,6)]
```






## Look at PUVI vs PALO


* There where two years where PALO was not visited, 2001 and 2002
* Neither site visited in 1999

```{r}
with(dat, table(Year,
                Location))
```


Equivalent
```{r}
table(dat$Location, dat$Year)
```


## Subset PUVI


```{r}
i.PUVI <- which(dat$Location == "PUVI")

dat.PUVI <- dat[i.PUVI,]

dim(dat)
dim(dat.PUVI)
```


### Use just species capture > 3 sep years


**ISSUE**: I think this is redundant - already done previously in prior script

```{r}
i4plus <- which(dat.PUVI$tot.years.obs > 3)

dat.PUVI <- dat.PUVI[i4plus,]

length(i4plus)
dim(dat.PUVI)

```



# GLMM set up

## Load library
```{r}
library(lme4)
```


## Set up for GLMM

GLMMs work best when continous covariates are centered.

```{r}
dat.PUVI$Year.cent <- with(dat.PUVI,
                               scale(Year,
                                     scale = F))
```


Poisson models are often overdispered; this can be corrected by including a observation level random effect, which can be made by giving each row of data  unique ID.

See 
Using observation-level random effects to model overdispersion in count data in ecology and evolution
XA Harrison - PeerJ, 2014 - peerj.com
https://peerj.com/articles/616/

```{r}
dat.PUVI$row.effect <- 1:dim(dat.PUVI)[1]
```

## Order disturbance sensitivty factor


Change from alphabetical order

```{r}
dat.PUVI$sensitivity <- factor(dat.PUVI$sensitivity,
                               levels = c("L",
                                          "M",
                                          "H"))
```


## Order migration "status" factor

Change from alphabetical order

```{r}
dat.PUVI$status <- factor(dat.PUVI$status,
                               levels = c("res",
                                          "NEO"))
```





# Explore data

Plot by disturbance sensitivity
```{r}
library(ggplot2)
library(cowplot)
qplot(y = spp.ann.tot,
      x = Year.cent,
      color = spp.code,
      data = dat.PUVI,
      facets = sensitivity ~ .)
```


# Explore data: migrants vs. residents

Plot by migration
```{r}
library(ggplot2)
library(cowplot)
qplot(y = spp.ann.tot,
      x = Year.cent,
      color = spp.code,
      data = dat.PUVI,
      facets = status ~ .)
```



# Fit a GLM to data to determine if trends are different between migrants and non migrants

A naive model would treat each data point as independent.  This is entirely inappropriat but used for illustration


## Null model: no trend
```{r}
glm.null <- glm(spp.ann.tot ~ 1,
data = dat.PUVI,
offset = log(net.hours),
family = poisson)
```
Plot by migration
```{r}
library(ggplot2)
library(cowplot)
qplot(y = spp.ann.tot,
      x = Year.cent,
      #color = spp.code,
      data = dat.PUVI) +
  geom_smooth(method = "lm")
```


## Annual trend
```{r}
glm.year <- glm(spp.ann.tot ~ Year.cent,
data = dat.PUVI,
offset = log(net.hours),
family = poisson)
```

## No annual trend, but different average abundance by status
```{r}
glm.status <- glm(spp.ann.tot ~ status,
data = dat.PUVI,
offset = log(net.hours),
family = poisson)
```




Plot by migration
```{r}
library(ggplot2)
library(cowplot)
qplot(y = spp.ann.tot,
      x = Year.cent,
      color = status,
      data = dat.PUVI) +
  geom_smooth(method = "lm")
```




## Trend, with different interactp

```{r}
glm.add <- glm(spp.ann.tot ~ Year.cent + status,
data = dat.PUVI,
offset = log(net.hours),
family = poisson)
```


## Trends vary by status

```{r}
glm.int <- glm(spp.ann.tot ~ Year.cent*status,
data = dat.PUVI,
offset = log(net.hours),
family = poisson)
```


## Compare models

These (inappropriate models) indicate that trends are different between migrants (NEO) and residents

```{r}
library(bbmle)

ICtab(type = "AICc",
      glm.null,
      glm.year,
      glm.status,
      glm.add,
      glm.int)
```


The slope fyr year*status is positive (and significant at p<0.05)
```{r}
summary(glm.int)
```


# Plot glm.year

ggplot can fit this model on the fly

```{r}
qplot(y = spp.ann.tot,
      x = Year.cent,
      #color = spp.code,
      data = dat.PUVI) + 
  #facet_grid(status~., scale = "free") +
 geom_smooth(method = "glm", se = FALSE,
              method.args = list(family = "poisson"))
```

## Fit a GLMM w/ an intercept for each species

* Species vary in their abundance; some are commone, some are rare
* We can have the model give each species its own intercept to account for its general level of abundance
* use lme4::glmer()
* add " + (1|Species) " 
* This is a "random intercepts"" model
* Assumes each species has its own intercept, and the distribution of intercepts can be modeld w/ a normal distribution.
* The model will fit a single slope and single intercept as its primary output
* Aroudn that intercept you can also determine the distribution of intercept terms


## Fit random intercepts model


```{r}
library(lme4)
rand.intercept <- glmer(spp.ann.tot ~ Year.cent +
      (1|spp.code),
      offset = log(net.hours),
data = dat.PUVI,
family = poisson)
```

### Model summary
```{r}
summary(rand.intercept)
```

### Compare to naive model

Slopes are similar, but intercepts are changed.

```{r}
rbind(glmm = fixef(rand.intercept),
      glm = coef(glm.year))
```


### Distribution of slopes

The slopes for each species
```{r}
ranef(rand.intercept)
```


Order by magnitude
```{r}
#extract random intercepts
my.ranefs <- ranef(rand.intercept)$spp.code

#put them in order
i <- order(my.ranefs[,1],decreasing = T)

my.ranefs$spp <- row.names(my.ranefs)


```

Look at relatively most abund and least abund
```{r}

rbind(head(my.ranefs[i ,])
      ,c("...","...")
,tail(my.ranefs[i ,]))
```


If we added the overall intercept we would get the mean abundance (on the log link scale).  

```{r}
mean.N.from.ranef <- exp(my.ranefs[i ,"(Intercept)"] + fixef(rand.intercept)[1])


mean.N.from.ranef <- cbind(data.frame(mean.N.from.ranef), my.ranefs$spp[i])

names(mean.N.from.ranef)[2] <- "spp.code"
```

We centered the year predictor; the intercepts therefore represent the abundance of these species the middle year of the study.  



The middle year is 2004
```{r}
with(dat.PUVI, table(Year, 
                     round(Year.cent)))
```


```{r}
i.2004 <- which(dat.PUVI$Year == 2004)
```


Average capture
```{r}
library(doBy)

mean.N.raw <- summaryBy(spp.ann.tot ~ spp.code, 
                        dat.PUVI)
```





Mean abundance in 2004 and slopes from random intercepts model are almost exactly the same

```{r}
temp <- merge(mean.N.raw,
              mean.N.from.ranef)


plot(temp$spp.ann.tot.mean,
     temp$mean.N.from.ranef)
abline(a = 0,b=1)
```





# Visualizing the random intercepts model


Generate predictions
```{r}
dat.PUVI$y.hat <- predict(rand.intercept,
                          type = "response")
```

Seperate line for each species
```{r}
randslopeplot <- qplot(y = spp.ann.tot,
      x = Year.cent,
      color = spp.code,
      data = dat.PUVI) +
  geom_line(aes(y = y.hat, 
                x = Year.cent,
                color = spp.code)) + theme(legend.position="none")


randslopeplot
```

NOTE: the lines DO have different slopes, BUT, this is b/c of the transformation from the Poisson models log link scale to ther eral scale.  The parameters themselves DO NOT have differents slopes for each species.s







# Random intercepts slopes

* What we want is a seperate model for each species
* AND and overall trend for the two "status" types
* This is a "random slopes" model
* It assumes that the slopes at the species-level are drawn from a normal distribution.
* Each species will also have its own intercept (drawn from a normal distribution)
* + (Year.cent|Species)
* "(Year.cent|spp.code)" fits sep intercept for each spp

```{r}
rand.slopes <- glmer(spp.ann.tot ~ Year.cent +
        (Year.cent|spp.code),
data = dat.PUVI,
offset = log(net.hours),
family = poisson)
```


```{r}

summary(rand.slopes)

```



# Visualizing the random slopes model




Generate predictions
```{r}
dat.PUVI$y.hat2 <- predict(rand.slopes,
                           type = "response")
```

Seperate line for each species
```{r}
randintplot <- qplot(y = spp.ann.tot,
      x = Year.cent,
      color = spp.code,
      data = dat.PUVI) +
  geom_line(aes(y = y.hat2, 
                x = Year.cent,
                color = spp.code)) + theme(legend.position="none")


randintplot
```




# compare the 2 models

```{r}
 library(cowplot)
plot_grid(randslopeplot, 
          randintplot)
```







# Include year random effect

* All measurements within a year are "nested" within a year
* They are not independent

```{r}
rand.slopes2 <- glmer(spp.ann.tot ~ Year.cent +
          (Year.cent|spp.code) +
          (1|Year),
data = dat.PUVI,
offset = log(net.hours),
family = poisson)

summary(rand.slopes2)
```


# Overdispersion


http://glmm.wikidot.com/faq

```{r}
overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
}
```



```{r}
overdisp_fun(rand.slopes3)
```




# Include observation level random effect

```{r}
rand.slopes3 <- glmer(spp.ann.tot ~ Year.cent +
            (Year.cent|spp.code) +
            (1|Year) +
            (1|row.effect),
data = dat.PUVI,
offset = log(net.hours),
family = poisson)

summary(rand.slopes3)
```






```{r}
overdisp_fun(rand.slopes3)
```



# Plot by species

```{r}


dat.PUVI$y.hat3 <- predict(rand.slopes3,
                           re.form = ~(Year.cent|spp.code),
type = "response")
```




Residents vs. migrants
```{r}
 qplot(y = spp.ann.tot,
      x = Year.cent,
      color = spp.code,
      data = dat.PUVI) +
  geom_line(aes(y = y.hat3, 
                x = Year.cent,
                color = spp.code)) +
  facet_grid(facets = .~status) +
  theme(legend.position="none")
```


```{r}
library(arm)
se.ranef(rand.slopes3)
```


```{r}
 temp <- qplot(y = spp.ann.tot,
      x = Year,
      color = status,
      data = dat.PUVI) +
  geom_line(aes(y = y.hat3, 
                x = Year,
                color = status)) +
  facet_wrap(facets =  ~spp.code  ,
             scales = "free") +
  theme(legend.position="none") +
  geom_vline(xintercept = 2003) +
  geom_vline(xintercept = 2002)

save_plot(temp,filename = "temp.jpeg",base_height = 10)
```











```{r}
rand.slopes3 <- glmer(spp.ann.tot ~ Year.cent +
            (Year.cent|spp.code) +
            (1|Year) +
            (1|row.effect),
data = dat.PUVI,
offset = log(net.hours),
family = poisson)

summary(rand.slopes3)
```


