---
title: "Lloyd 2016 PeerJ Dominican Birds: Data Analysis"
author: "brouwern@gmail.com"
date: "February 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Re-Analysis of Lloyd 2016 data


## Load Lloyd data

```{r}
spp.ann.tot2 <- read.csv(file = "Lloyd_PeerJ2916_cleaned_data_2_7_2017.csv")
```

## Look at data

```{r}
summary(spp.ann.tot2)
dim(spp.ann.tot2)
```


## Hard code data transformations

### Captures/net.hours.1K

"For the purposes of standardization with other constant-effort mist-netting studies, we report capture rate per 1,000 net hours (i.e., expected captures for every 1,000 h that 12-m net was open)." (page 4)

```{r}
spp.ann.tot2$caps.per.1K.nethours <- with(spp.ann.tot2, spp.ann.tot/net.hours * 1000)
```


### log(net.hours)

'We accounted for variation in capture effort by using the number of net hours per capture session (log-transformed) as an offset in the model." (page 4)

```{r}
spp.ann.tot2$net.hours.log <- log(spp.ann.tot2$net.hours)
```






## Subset for species 

### Indicies
```{r}
# Rufous-throated Solitaire (Myadestes genibarbis)
i.RTSO <- which(spp.ann.tot2$Species ==  "RTSO")

#Green-tailed Ground-Tanager (Microligea palustris)
i.GTGT <- which(spp.ann.tot2$Species ==  "GTGT")

#Greater Antillean Bullfinch (Loxigilla violacea)
i.GABU <- which(spp.ann.tot2$Species ==  "GABU")

#Western Chat-Tanager (Calyptophilus tertius)
i.WCHT <- which(spp.ann.tot2$Species ==  "WCHT")

#Hispaniolan Highland-Tanager (Xenoligea montana)
i.HHTA <- which(spp.ann.tot2$Species ==  "HHTA")
```
 
 
### Dat subset

```{r}
rtso <- spp.ann.tot2[i.RTSO,]
gtgt <- spp.ann.tot2[i.GTGT,]
gabu <- spp.ann.tot2[i.GABU,]
wcht <- spp.ann.tot2[i.WCHT,]
hhta <- spp.ann.tot2[i.HHTA,]
```

<br>




# Plot raw data with ggplot

## Figure 1:

* I'm not sure exactly how they made their plot, since the y plot captures/net.hours*1000, but model captures w/log(net.hours) offset

```{r}

library(ggplot2)
qplot(y =caps.per.1K.nethours,
      x = Year,
      data = rtso) +
  facet_grid(Species ~ Location, scale  = "free") +
  geom_smooth(method = "glm", 
              method.args = list(family = "poisson"))
```

## Figure 1 caption
"Figure 1 Trends in capture rate of Rufous-throated Solitaire (Myadestes genibarbis). Observed cap- ture rate (dots) of Rufous-throated Solitaire and changes in expected capture rate (solid line; shaded area =95% confidence interval) per 1,000 net-hours at two sites (PALO, Palo de Agua; PUVI, Pueblo Viejo) in montane broadleaf forest of Sierra de Bahoruco, Dominican Republic."


<br>

# Model data

## Notes on modeling

Their methods:
"The predictor variables included year, site, and the interaction between site and year. We considered three models for each species: year only, site+year, and site???year"

For the sake of completeness I'll fit some additional mdoels


## Results

* "Capture rates declined over the course of our study for Rufous-throated Solitaire (beta.year = ???0.04, 95% CL= ???0.076,???0.001; P = 0.04; Fig. 1)"
* "Green-tailed Ground- Tanager (beta.year = ???0.07; 95% CL= -0.092,???0.040; P < 0.001; Fig. 2)"
* Greater Antillean Bullfinch (beta.year =???0.05; 95% CL=???0.097,???0.009, P =0.02; Fig. 3).


## Fit models

* glm() instead of lm()
* "~ 1" fits null model
* "family = poisson "
* "offset = net.hours.log" applies correct for variable effort between years

### Fit null model: NO change over time

* No change over time (no slope)
* AND no difference between sites (single intercept)
* They did not fit this model

```{r}
rtso.null<- glm(spp.ann.tot ~ 1,
            family = poisson,
            offset = net.hours.log,
            data = rtso)
```

### Fit Location model

* No chagne over time (no slope)
* Different mean abundance for each site (different intercepts)
* One site gets set as the basline and labeled the "INtercept"
    + Goes in alphabetical order unless you recode things
    + So PALO is the baseline
    + This is on the Poisson link scale
    + need to be transformed
* The 2nd site is defined as a difference between teh baseline PALO and the 2nd site PUVI.
    + The model outputs "LocationPUVI".
    + This is a difference, not an actual intercept
    + "(Intercept)"+"LocationPUVI" = PUVIs mean abundance
    + Again, on Poisson link scale

```{r}
rtso.Loc <- glm(spp.ann.tot ~ Location,
            family = poisson,
            offset = net.hours.log,
            data = rtso)
```



### Fit year model

* This was their simplest model

```{r}
rtso.Year<- glm(spp.ann.tot ~ Year,
            family = poisson,
            offset = net.hours.log,
            data = rtso)
```


### Fit model with year + location

* I often call this an "additive" model, in contrast to a "multiplicative" model which contains an interaction
* This was their best supported model for most/all species

```{r}
rtso.add<- glm(spp.ann.tot ~ Year + Location,
            family = poisson,
            offset = net.hours.log,
            data = rtso)
```




### Fit model with year + location

* I often call this an "multiplicative" model, b/c it contains an interaction term
* It would also becalled a "full" model b/c it contain all of the terms being considered
* They also considered models w/ a parameter for intensity of El Nino, but I don't have those data yet.
* NOt that the 2 versions are equivalent; R can expand out the equation on it wn 
```{r}
rtso.mult<- glm(spp.ann.tot ~ Year + Location + Year*Location,
            family = poisson,
            offset = net.hours.log,
            data = rtso)


rtso.mult<- glm(spp.ann.tot ~ Year*Location,
            family = poisson,
            offset = net.hours.log,
            data = rtso)
```



# Get confidence itnervals for parameters
```{r}
confint(rtso.add)
```


# Compare all models w/AIC

They did not do this

```{r}
library(bbmle)
AICtab(rtso.null,rtso.Loc,rtso.Year,rtso.add,rtso.mult)
```


