---
title: "GLM loop"
author: "brouwern@gmail.com"
date: "May 24, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load data
```{r}
#Main direotry for lloyd project
d.working <- "/avian_ecology_replications/Lloyd_2016_PeerJ_Dominican_birds/"

#data directory
d2 <- paste(d.working,
            "data/data_out", sep = "")

list.files(path = paste(".", d2,sep = ""))

working.dat2 <- read.my.csv(rel.dir. = d2,
            file. = "Lloyd_data_cleaned_with_meta")       
```


## Calc Captures per 1000 net hours

```{r}
working.dat2$caps.per.1K.nethours <- with(working.dat2,
                                         spp.ann.tot/net.hours*1000)
```


## Calc log net hours

In case "Inf" is returned, clean it up; this is no longer and issue and was likely due to an error in data processing.

```{r}
working.dat2$net.hours.log <- log(working.dat2$net.hours)

i.INF <- which(working.dat2$net.hours.log == -Inf)

working.dat2$net.hours.log[i.INF] <- NA

```



## Focal species

11 Focal species in original Lloyd et al 2016 paper.  Some have figures, others only discussed in text.
```{r}
focal.spp <- c("NBTO",
               "RTSO",
               "BITH",
               "OVEN",
               "BTBW",
               "HHTA",
               "GTGT",
               "BCPT",
               "WCHT",
               "GABU",
               "HISP")

length(focal.spp)
```



## Loop over species list


```{r}
#Initialize list
list.focal.spp.mods <- list()


for(i in 1:length(focal.spp)){
  i.use <- which(working.dat2$Species == focal.spp[i])
  
  print(i)
  print(focal.spp[i])
  #print(i.use)
  
  
  # Models
  ## null model - intercept only
  mod.null <- glm(spp.ann.tot ~ 1,
            family = poisson,
            offset = net.hours.log,
            data = working.dat2[i.use,])
  
  ## year only
  mod.yr   <-update(mod.null, .~. + Year)
  
  ## Location only PALO vs. PUVI 
  mod.loc  <-update(mod.null, .~. +        Location)
  
  ## Additive model
  mod.add  <-update(mod.null, .~. + Year + Location)
  
  ## Multiplicative mdoel
  mod.mult <-update(mod.null, .~. +               Year*Location)
  
  ## compile list of models
  list.focal.spp.mods[[i]] <- list(null = mod.null,
                                   yr   = mod.yr,
                                   loc  = mod.loc,
                                   add  = mod.add,
                                   mult = mod.mult)
  
  ## add names
  names(list.focal.spp.mods)[[i]] <- paste(focal.spp[i],
                                           "poisson",
                                           sep = ".")
  
}
```



## Build table of models

For each type of model (year only, year + site, etc) I build a table for the species used and then combine them at the end


### year only model
Dataframe for storage

```{r}
df.yr <- expand.grid(spp = focal.spp,
                  NLB.int = NA,
                  NLB.yr = NA)
```

Loop over species to get betas
```{r}
for(i in 1:length(focal.spp)){
  df.yr[i,-1] <- coef(list.focal.spp.mods[[i]]$yr)
}


```


## year + site model

Dataframe for storage
```{r}
df.add <- expand.grid(spp = focal.spp,
                  NLB.int = NA,
                  NLB.yr = NA,
                  NLB.site.PUV = NA) #categorical variable for site
```



Loop over species to get betas
```{r}
for(i in 1:length(focal.spp)){
  df.add[i,-1] <- coef(list.focal.spp.mods[[i]]$add)
}


```




## year*site model
Dataframe for storage

```{r}
df.mult <- expand.grid(spp = focal.spp,
                  NLB.int = NA,
                  NLB.yr = NA,
                  NLB.site.PUV = NA,#categorical variable for site
                  NLB.siteXloc = NA)#year x site interaction
```



Loop over species to get betas for year*site model

```{r}
for(i in 1:length(focal.spp)){
  df.mult[i,-1] <- coef(list.focal.spp.mods[[i]]$mult)
}
```




## Combine models output

```{r}
#annotate with type of model
df.yr$mod <- "year"
df.add$mod <- "add"
df.mult$mod <- "mult"

#add empty columns so all of the dfs match
df.yr$NLB.site.PUV <- NA
df.yr$NLB.siteXloc <- NA

df.add$NLB.siteXloc <- NA

#stack dataframes
df2 <- rbind(df.yr,
             df.add,
             df.mult)

```




### Check against published beta

Load Lloyd Table 2
```{r}
f. <- paste(".",
      d.working,
      "tables/Lloyd_coefs_vs_our_coefs/",
      "Lloyd_regression_coefs.csv",
      sep = "")


orig.reg.coefs <- read.csv(file =f.)

orig.reg.coefs$modeler <- "Lloyd"
```



```{r}
options(digits=2)
df2$modeler <- "NLB"
df3 <- merge(orig.reg.coefs,
             df2,all = T)

```


### Compare outputs

In general my estiamtes are very similar, indicating that I can reproduce their data cleaning workflow and analysis.  However, for 1 species there is a major discrepancy and for 2 more there are minor ones.

* BITH - major discrepancy
* GTGT - their estimate is 0.01 lower
* HISP, OVEN = their estimate is 0.001

There are other discrepancies of similar magnitude (0.001 to 0.01) where my estiamtes is lower but these are consistent with them rounding up their parametrs to nearest whole 0.01 or 0.001; that is, I can round my estimate up to get their value.  I consider it more of a problem when my estimate is higher, since their rounding produced an estimate that still wasn't simialr to mine.

```{r}
df3[,c("spp","Fig","mod",
       "B.yr","NLB.yr",
       "B.site","NLB.site.PUV")]
```



