---
title: "Lloyd 2016 DR Birds: Both Locations GLMM"
author: "brouwern@gmail.com"
date: "February 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Libraries

```{r}
library(lme4)
library(bbmle)
```


##Data setup

### Load data
```{r}
#Main direotry for lloyd project
d.working <- "/avian_ecology_replications/Lloyd_2016_PeerJ_Dominican_birds/"

#data directory
d2 <- paste(d.working,
            "data/data_out", sep = "")

list.files(path = paste(".", d2,sep = ""))

working.dat2 <- read.my.csv(rel.dir. = d2,
            file. = "Lloyd_data_cleaned_with_meta")   

```


Log the net hours
```{r}
working.working.dat2$net.hours.log <- log(working.working.dat2$net.hours)
```



### Load spp info

Not done

* skip = 7 skips the 7 rows of header info in this field
```{r}
    # spp.info <- read.csv(file = "Latta_Lloyd_species_list_SLvs1.csv",
    #                      skip = 7)

```

#### Merge

```{r}
    # names(working.dat2)[3] <-"spp.code"
    # 
    # summary(working.dat2$spp.code)
    # summary(spp.info$spp.code)
    # 
    # 
    # dat2 <- merge(working.working.dat2, spp.info, by =)
    # 
    # dim(dat2)
    # dim(working.working.dat2)
    # 
    # working.working.dat2 <- dat2
    # 
    # 
    # #remove some extra rows
    # working.working.dat2 <- working.working.dat2[,-c(2,4,6)]
```


* Dataframe should be ~547 rows




### Data subset

Use just species capture > 3 sep years, since you need 3+ points to fit a line.


```{r}
i4plus <- which(working.dat2$yrs.obs > 3)

working.dat3 <- working.dat2[i4plus,]
 
dim(working.dat2)
dim(working.dat3)
```



# GLMM set up

## Set up for GLMM

GLMMs work best when continous covariates are centered.

```{r}
working.dat2$yr.cnt <- with(working.dat2,
                               scale(Year,
                                     scale = F))
```


Poisson models are often overdispered; this can be corrected by including a observation level random effect, which can be made by giving each row of data  unique ID.

See 
Using observation-level random effects to model overdispersion in count data in ecology and evolution
XA Harrison - PeerJ, 2014 - peerj.com
https://peerj.com/articles/616/

```{r}
working.dat2$row.effect <- 1:dim(working.dat2)[1]
```

## Order disturbance sensitivty factor

```{r}
    # working.dat2$sensitivity <- factor(working.dat2$sensitivity,
    #                                levels = c("L","M","H"))
```


## Order migration "status" factor

```{r}
    # working.dat2$status <- factor(working.dat2$status,
    #                                levels = c("res","NEO"))
```


## Overdispersion


http://glmm.wikidot.com/faq
```{r}
overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```







## Model data

### Models w/ random slopes

```{r}
m.ranefs.base <- glmer(spp.ann.tot ~ 1 +
            (1|Year:Location) +
            (1|Year) +
            (1|row.effect),
data = working.dat2,
offset = log(net.hours),
family = poisson)



m.ranslope.spp <- update(m.ranefs.base, . ~ . + 
                           (     1|Species:Location) +
                           (yr.cnt|Species) )
  

m.ranslope.spp.x.site <- update(m.ranefs.base, . ~ . + 
                           (yr.cnt|Species:Location) +
                           (     1|Species) )




```



### Compare models

```{r}

bbmle::AICtab(m.ranefs.base,
              m.ranslope.spp,
              m.ranslope.spp.x.site)
```






















## Include observation level random effect


```{r}
m1.no.Loc <- glmer(spp.ann.tot ~ 
                        yr.cnt +
            (yr.cnt|Species) +
            (1|Year) +
            (1|row.effect),
data = working.dat2,
offset = log(net.hours),
family = poisson)

summary(m1.no.Loc)
```





```{r}
working.dat2$log.net.hours <- with(working.dat2, log(net.hours))
m1.Loc.ranef <- lmer(spp.ann.tot/log.net.hours ~ 
                        yr.cnt +
                     (1|spp.code) +
            (yr.cnt|spp.code:Location) +
            (1|Year) +
            (1|Year:Location), #+
            #(1|row.effect),
data = working.dat2#,
#offset  = log.net.hours,
#family = poisson
#
)

summary(m1.Loc.ranef)
```


# Predictions

expand.grid(
  spp.code = levels(m1.Loc.ranef@frame$spp.code),
  Location = levels(m1.Loc.ranef@frame$Location),
  yr.cnt = range(m1.Loc.ranef@frame$yr.cnt),
  Year = unique(m1.Loc.ranef@frame$Year)
  net.hours = 1000
)


```{r}

fake.working.dat2 <- m1.Loc.ranef@frame

summary(fake.working.dat2)
summary(log(working.dat2$net.hours))


#fake.working.dat2$"(offset)" <- 1

y.hat <- predict(m1.Loc.ranef,
                 newdata =fake.working.dat2,
                re.form = ~ (1|spp.code) + (yr.cnt|spp.code:Location))


summary(y.hat)


out <- m1.Loc.ranef@frame
out$y.hat <- y.hat


summary(out$y.hat)

library(ggplot2)
library(cowplot)

out$spp.ann.tot/exp(out$`(offset)`)*1000

temp <- qplot(y = spp.ann.tot/log.net.hours,
      x = Year,
      color = Location,
      data = out) +
  geom_line(aes(y = y.hat, 
                x = Year,
                color = Location)) +
  facet_wrap(facets =  ~spp.code  ,
             scales = "free") +
  theme(legend.position="none") #+
  #geom_vline(xintercept = 2003) +
  #geom_vline(xintercept = 2002)

save_plot(temp,filename = "temp.jpeg",base_height = 10)


```







```{r}

```




```{r}
library(arm)
se.list <- se.ranef(m1.Loc.ranef)
str(se.list,1)

se.list[[2]][,"yr.cnt"]
```






```{r}
ranef(m1.Loc.ranef)[2]
```












```{r}
rand.slopes3 <- glmer(spp.ann.tot ~ yr.cnt +
            (yr.cnt|spp.code) +
            (1|Year) +
            (1|row.effect),
data = working.dat2,
offset = log(net.hours),
family = poisson)

summary(rand.slopes3)
```



