---
title: "Lloyd sensitivity analysis"
author: "Emily Scott"
date: "September 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PART 0: Set initial values
```{r}
### Minimum number of species to include in mixed model
min.spp <- 4

### Selection criteria (year)
### Minimum number of years captured to be included in model
min.yr <- 5

### Selection criteria (individuals)
### Minimum number of inidividuals captured to be included in model
min.id <- 4

### Do you want to use minimum number of years captured, or minimum number of individuals captured as selection criteria?
### Options: "year" or "id"
select.crit <- "year"
```



# PART 1: Load packages, set working directory, load data in long and wide formats
## Load packages
```{r}
library(lme4)
library(reshape2)
library(ggplot2)
library(cowplot)
library(reshape2)
```

## Set working directory and load long-format data
```{r}
setwd("C:/Users/emily/OneDrive/Documents/R/git/Aviary/replications/avian_ecology_replications/Lloyd_2016_PeerJ_Dominican_birds")
lloyd.dat.long <- read.csv("./data/data_out/Lloyd_data_cleaned.csv")
```



# PART 2: Prep data for mixed model
## Remove second session data
```{r}
lloyd.dat.long <- lloyd.dat.long[-(which(lloyd.dat.long$Session == 2)),]
```

## Create dataframe with data in wide format
```{r}
lloyd.dat.wide <- dcast(lloyd.dat.long, Species + Location ~ Year, value.var = "spp.ann.tot")
```

## Create list of species codes
```{r}
spp.list <- list() 
spp.list <- unique(as.character(lloyd.dat.wide$Species))
spp.list <- as.list(spp.list)
```

## Create index for each species
```{r}
for (i in 1:length(spp.list))
{
    assign(paste("i", spp.list[i], sep = "."), which(lloyd.dat.long$Species == spp.list[i]))
}
```

## Replace NAs with 0s so selection criteria works correctly
```{r}
i.NA <- is.na(lloyd.dat.wide[,5] == TRUE)
lloyd.dat.wide[i.NA,5] <- 0

i.NA <- is.na(lloyd.dat.wide[,6] == TRUE)
lloyd.dat.wide[i.NA,6] <- 0
```

## Reshape
## Find species that don't meet selection criteria
```{r}
### This chunk runs if the selection criteria is set to number of years captured
if (select.crit == "year")
{
  ### Create column with total number of years captured
  lloyd.dat.wide$yr.cap <- rowSums(lloyd.dat.wide[,(3:15)] != 0)
  ### Index species with insufficient captures for selection criteria
  i.drop1 <- which(lloyd.dat.wide$yr.cap < min.yr)
  ### Create list of species codes corresponding to species with too few individuals captured
  bad.spp <- lloyd.dat.wide[i.drop1,c("Species","Location")]
  ### Index rows in dataframe that correspond to species that didn't meet selection criteria
  i.drop2 <- which(lloyd.dat.long$Species %in% bad.spp[,1] & lloyd.dat.long$Location %in% bad.spp[,2])
  ### Create working dataframe
  dat.working <- lloyd.dat.long[-i.drop2, ]
}

### This chunk runs if the selection criteria is set to number of individuals captured
if (select.crit == "id")
{
  ### Create column with total number of individuals captured
  lloyd.dat.wide$id.cap <- rowSums(lloyd.dat.wide[,(3:15)])
  ### Index species with insufficient captures for selection criteria
  i.drop1 <- which(lloyd.dat.wide$id.cap < min.id)
  ### Create list of species codes corresponding to species with too few individuals captured
  bad.spp <- lloyd.dat.wide[i.drop1, c("Species","Location")]
  ### Index rows in dataframe that correspond to species that didn't meet selection criteria
  i.drop2 <- which(lloyd.dat.long$Species %in% bad.spp[,1] & lloyd.dat.long$Location %in% bad.spp[,2])
  ### Create working dataframe
  dat.working <- lloyd.dat.long[-i.drop2, ]
}

### Create working species list by removing species that didn't meet selection criteria
spp.list.working <- list()
spp.list.working <- unique(as.character(dat.working$Species))
spp.list.working <- as.list(spp.list.working)
```

## Add column for centered year to working dataframe (i.e. center year around 0)
```{r}
dat.working$yr.cnt <- scale(dat.working$Year, scale = F)

dat.working$log.net.hours <- with(dat.working, log(net.hours))
```


# PART 3: Prep data for sensitivity analysis
## Sort species by number of years/individuals captured (set by selection criteria)
```{r}
if (select.crit == "year")
{
  sorted.df <- lloyd.dat.wide[order(lloyd.dat.wide$yr.cap), ]
  i.keep <- which(colnames(lloyd.dat.wide) %in% c("Species", "Location", "yr.cap"))
  sorted.df <- sorted.df[,i.keep]
}

if (select.crit == "id")
{
  sorted.df <- lloyd.dat.wide[order(lloyd.dat.wide$id.cap), ]
  i.keep <- which(colnames(lloyd.dat.wide) %in% c("Species", "Location", "id.cap"))
  sorted.df <- sorted.df[,i.keep]
}
```

## Drop sorted species with fewest caps (same species as were dropped when creating dat.working above)
```{r}
### Index species to be removed 
i.drop3 <- which(sorted.df$Species %in% bad.spp[,1] & sorted.df$Location %in% bad.spp[,2])

### Remove species
sorted.df <- sorted.df[-i.drop3,]
```

## Create matrix to store slopes
```{r}
### Want a row for every species 
### Want a column for each mixed model (4 less than total number of species because mixed model doesn't behave with <5 species)
mixed.model.summary <- matrix(data = NA, nrow = length(spp.list.working), ncol = 1)
for (i in 1:length(spp.list.working))
{
  mixed.model.summary[i] <- spp.list.working[[i]]
}
mixed.model.summary <- as.data.frame(mixed.model.summary)
colnames(mixed.model.summary) <- c("Species")
```

## Run mixed model with fewer species each time 
```{r}
### Clone dat.working so it doesn't get overwritten during looping process
updated.dat.working <- dat.working

### Loop through (min.spp) times less than the total number of species because mixed model doesn't behave well with too few species
for (i in 1:c(length(sorted.df[,2])-min.spp))
{
  ### Mixed model
 
  m1.Loc.ranef <- lmer(spp.ann.tot/log.net.hours ~ 1 + yr.cnt +
                          (1|Species) +
                          (yr.cnt|Species:Location) +
                          (1|Year) +
                          (1|Year:Location),
                        data = updated.dat.working)
  
  ### Store current coefficients
  current.coeff <- coef(m1.Loc.ranef)$Species[,1]
  
  ### Convert coefficients to dataframe
  current.coeff <- as.data.frame(current.coeff)
  
  ### Add column containing species code (for the purpose of merging)
  current.coeff$Species <- rownames(coef(m1.Loc.ranef)$Species)
  
  ### Merge current coefficients with mixed model summary
  mixed.model.summary <- merge(mixed.model.summary, current.coeff, by = c("Species"), all.x = TRUE)
  
  ### Rename column
  colnames(mixed.model.summary)[i+1] <- paste("M", i, sep = ".")

  # Delete next least abundant species
  i.drop <- which(updated.dat.working$Species == sorted.df$Species[i] & updated.dat.working$Location == sorted.df$Location[i])
  updated.dat.working <- updated.dat.working[-i.drop,]
}
```



# PART 4: Plotting
## Prepare dataframe for plotting
```{r}
### Change to dataframe
mixed.model.summary <- as.data.frame(mixed.model.summary)

### Reshape data to long format
dat.melt2 <- melt(data = mixed.model.summary, id.vars = "Species")
```

## Plot
```{r}
### Visualize species-specific slope coefficients (use qplot)
qplot(y = value,
      x = variable,
      data = dat.melt2,
      color = Species,
      group = Species,
      geom = c("line"),
      xlab = "Model",
      ylab = "Parameter estimate") 
```

## Plot
```{r}
### Visualize species-specific slope coefficients (use ggplot)
ggplot(data = dat.melt2,
       aes(x = variable, y = value, colour = Species, group = Species)) +
  geom_line(size = 1.0) +
  background_grid(major = "xy", minor = "none") +
  ggtitle("Sensitivity Analysis") +
  xlab("Model") +
  ylab("Parameter estimate")
```




